\huge The human visual system preserves the hierarchy of 2-dimensional pattern regularity

### 2021-02-24 Writeup/kohler_clarke_2021.tex
W literature added

We observe a reliable correlation between our brain imaging and psychophysical data. This suggests that the two measurements reflect the same underlying symmetry representations in visual cortex. While it should be noted that the correlation is relatively modest ($R^2 = 0.44$), we note that this may be partly due to the fact that the same individuals did not participate in the two experiments. Future work in which behavioral and brain imaging data are collected from the same participants, will help further establish the connection between the two measurements, and tease apart any additional complexity that may not have been captured by the summary statistics we applied here.

We also find an of index for both our brain imaging measurements and our symmetry detection thresholds. This means that the visual system not only represents the hierarchical relationship captured by individual subgroups, but also distinguishes between subgroups depending on how many times the subgroup is repeated in the supergroup, with more repetitions leading to larger pairwise differences. Our measured effect of index is relatively weak. This is perhaps because the index analysis does not take into account the \textit{type} of isometries that differentiate the subgroup and supergroup. The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure 1. The two groups are comparable except \textit{PM} contains reflection and \textit{PG} contains glide reflection, and the former clearly elicits higher amplitudes and lower thresholds. The charting of how different symmetry types contribute to the representational hierarchy will be an important goal for future work.

>>>>

We observe a reliable correlation between our brain imaging and psychophysical data. This suggests that the two measurements reflect the same underlying symmetry representations in visual cortex. While it should be noted that the correlation is relatively modest ($R^2 = 0.44$), we note that this may be partly due to the fact that the same individuals did not participate in the two experiments. Future work in which behavioral and brain imaging data are collected from the same participants, will help further establish the connection between the two measurements, and tease apart any additional complexity that may not have been captured by the summary statistics we applied here. It has recently been demonstrated that \textit{W}, a measure of perceptual goodness derived from a holographic model of regularity \citep{van_der_helm_1996}, can predict EEG responses \citep{makin_w_2016} and perceptual discrimination performance \citep{nucci_2007} for patterns that contain symmetry and other types of regularity. The model was formulated based on dot patterns with symmetry axes centered on a single spatial location. It will important to determine if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane, and test how well it can explain behavioral and brain responses to wallpapers.

We also find an effect of index for both our brain imaging measurements and our symmetry detection thresholds. This means that the visual system not only represents the hierarchical relationship captured by individual subgroups, but also distinguishes between subgroups depending on how many times the subgroup is repeated in the supergroup, with more repetitions leading to larger pairwise differences. Our measured effect of index is relatively weak. This is perhaps because the index analysis does not take into account the \textit{type} of isometries that differentiate the subgroup and supergroup. The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure 1. The two groups are comparable except \textit{PM} contains reflection and \textit{PG} contains glide reflection, and the former clearly elicits higher amplitudes and lower thresholds. An important goal for future work will be to map out how different symmetry types contribute to the representational hierarchy.
